---
title: "PA 9.1: Mystery Animal"
subtitle: "Linear Regression"
format: 
  html:
    embed-resources: true
    code-tools: true
    toc: true
    number-sections: true
editor: source
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

```{r setup}
#| include: false
library(tidyverse)
```

[Download starter .qmd file](PA9.2-mystery-animal.qmd)

## Data

The data set contains the weights of a particular animal before and after a year of eating only roasted duck.

```{r}
mystery_animal <- read_csv("https://github.com/earobinson95/stat331-calpoly/raw/master/practice-activities/data/mystery_animal.csv")
```

## Linear Regression

Fit a linear regression to determine if the Duck Diet is associated with the animal gaining weight, losing weight, or neither.

```{r}
# Fit linear regression model here
```

## Residuals

Extract and plot the residuals *(observed value minus predicted value)* of your linear model versus `weight_before`.

::: callout-tip
There are a few different ways to obtain your residuals. My favorite is the `augment()` function from the {broom} package. I like this option a lot because it gives you all of the information from your linear regression in a tidy [tibble]{.underline}!

Each row has the explanatory and response values, the fitted / predicted value (`.fitted`), and the residual (`.resid`).
:::

```{r}
# Extract and plot the residuals here
```

## Canvas Submission: Mystery Animal

> What animal do you see?
